---
title: "structure_analysis"
author: "Amanda Stahlke"
date: "2/28/2020"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(pophelper)
library(gridExtra)
library(gtable)
library(label.switching)

cbbPalette <- c("#56B4E9", "#E69F00", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
human_colors <- c('blue', 'orange', 'green', 'yellow', 'dark blue', 'red', 'pink')
spp <- c('carinata', 'elongata', 'sublineata', 'carinulata', 'elongata_sub', 'carinulata_sub', 'other')

legend_colors <- as.data.frame(cbind(cbbPalette, human_colors, spp), stringsAsFactors = F)
legend_colors

```

## Structure 

Examined "global" ancestry for all available samples. Old replicates and some suspected contamination/switching have been removed. The masterpopmap has ALL of these samples. 

Admixture model, LD=0, allele freqs are uncorrelated, alpha is inferred for each popualtion.

```{r inputs}
masterpopmap <- read.delim("../info/masterpopmap.tsv",
                  sep = "\t",
                   col.names = c("sampID", "popID"),
                   stringsAsFactors = FALSE, header = FALSE)

sfiles <- list.files(path="~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/Results/", 
  pattern=glob2rx("outfile_*f"),full.names=TRUE)

slist <- readQ(sfiles,filetype="structure", 
               indlabfromfile = TRUE, 
               readci = TRUE)

popmap <- left_join(data.frame(sampID = rownames(slist[[1]])), 
                    masterpopmap)

popID <- data.frame(popID = popmap$popID)
popID <- data.frame(lapply(popID, as.character), stringsAsFactors=FALSE)
sapply(popID, is.character)

sample_meta.dat <- read.csv("~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/info/pop_metadata.csv", header = TRUE, stringsAsFactors = FALSE)
sample_meta.dat

pop_order <- sample_meta.dat[,c("popID", "Order")]
pop_order <- arrange(pop_order, Order)[,1]
pop_order
```


```{r}
samplesize <- left_join(popmap, sample_meta.dat) %>%
  count(popID) %>%
  merge(sample_meta.dat)

write.table(samplesize, "pop_metadata_samplesize.tsv", sep = "\t", row.names = F, quote = F)

```

## What is the best K?

```{r which_K}
sr1 <- summariseQ(tabulateQ(slist))
p <- evannoMethodStructure(data=sr1,exportplot=F,returnplot=T,returndata=F,basesize=12,linesize=0.7)
grid.arrange(p)

evannoMethodStructure(data=sr1,
                  exportplot=T,
                  exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/plots/",
                outputfilename = "global_deltaK",
                returndata=F,
                height = 10, width = 9)
```

K=2 and K=3 are super stable. D. carinata appears as a mix between sublineata and elongata. I am surprised to see, for the first time in these samples, indications of admixture between carinlata and sublineata along the Rio Grande. 

```{r Ks2-3}
k2 <- plotQ(alignK(slist[c(21:30)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            subsetgrp = pop_order,
            ordergrp = T,
            grplab = popID,
            linepos = 1,
            grplabangle=-45,
            grplabpos = 0, 
            grplabheight=20)
grid.arrange(k2$plot[[1]])

k3 <- plotQ(alignK(slist[c(31:40)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k3$plot[[1]])
```
As usual, K=4 reflects lab populations, elongata in the native range, and hints at carinulata substructure. 

```{r alignK4-8}

k4 <- plotQ(alignK(slist[c(41:50)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            grplab = popID,
            ordergrp = T,
            grplabangle=-90)
grid.arrange(k4$plot[[1]])


k5 <- plotQ(alignK(slist[c(51:60)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            grplab = popID,
            ordergrp = T,
            grplabangle=-90)
grid.arrange(k5$plot[[1]])

k6 <- plotQ(alignK(slist[c(61:70)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            grplab = popID,
            ordergrp = T,
            grplabangle=-90)
grid.arrange(k6$plot[[1]])


k7 <- plotQ(alignK(slist[c(71:80)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k7$plot[[1]])

k8 <- plotQ(alignK(slist[c(81:90)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            # clustercol=cbbPalette,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k8$plot[[1]])

```


```{r K4}
plotQMultiline(slist[41],
      showticks=T,
      clustercol=cbbPalette,
      grplab = popID,
      ordergrp=TRUE,
      spl = 25,
      showindlab = T,
      showyaxis = T,
      barsize = .9,
      subsetgrp = pop_order,
      basesize=11, 
      exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/plots/",
      outputfilename = "multilineK4",
      useindlab = T)

```

```{r plots2save_repsK}
plotQ(alignK(slist[c(21:30)]),
            imgoutput="join",
                  exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/plots/",
  outputfilename =  "Joined-K2",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=cbbPalette,
  subsetgrp = pop_order,
  height = .1,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  grplabangle=45,
  grplabpos = .8,
  grplabspacer = 0,
  grplabheight=70,
  width=100,
  showtitle=T,
  grplabsize = 4,
  titlelab="Global Population Structure, K=2",
  titlesize = 20)

plotQ(alignK(slist[c(31:40)]),
            imgoutput="join",
                  exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/plots/",
  outputfilename =  "Joined-K3",    
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=cbbPalette,
  subsetgrp = pop_order,
  height = .1,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  grplabangle=45,
  grplabpos = .8,
  grplabspacer = 0,
  grplabheight=70,
  width=100,
  showtitle=T,
  grplabsize = 4,
  titlelab="Global Population Structure, K=3",
  titlesize = 20)

plotQ(alignK(slist[c(41:50)]),
            imgoutput="join",
            exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/plots/",
  outputfilename =  "Joined-K4",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=cbbPalette,
  subsetgrp = pop_order,
  height = .1,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  grplabangle=45,
  grplabpos = .8,
  grplabspacer = 0,
  grplabheight=70,
  width=100,
  showtitle=T,
  grplabsize = 4,
  titlelab="Global Population Structure, K=4",
  titlesize = 20)

plotQ(alignK(slist[c(51:60)]),
            imgoutput="join",
            exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/plots/",
  outputfilename =  "Joined-K5",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=cbbPalette,
  subsetgrp = pop_order,
  height = .1,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  grplabangle=45,
  grplabpos = .8,
  grplabspacer = 0,
  grplabheight=70,
  width=100,
  showtitle=T,
  grplabsize = 4,
  titlelab="Global Population Structure, K=5",
  titlesize = 20)

plotQ(alignK(slist[c(61:70)]),
            imgoutput="join",
            exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/plots/",
  outputfilename =  "Joined-K6",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=cbbPalette,
  subsetgrp = pop_order,
  height = .1,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  grplabangle=45,
  grplabpos = .8,
  grplabspacer = 0,
  grplabheight=70,
  width=100,
  showtitle=T,
  grplabsize = 4,
  titlelab="Global Population Structure, K=6",
  titlesize = 20)

plotQ(alignK(slist[c(71:80)]),
            imgoutput="join",
            exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/plots/",
  outputfilename =  "Joined-K7",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=cbbPalette,
  subsetgrp = pop_order,
  height = .1,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  grplabangle=45,
  grplabpos = .8,
  grplabspacer = 0,
  grplabheight=70,
  width=100,
  showtitle=T,
  grplabsize = 4,
  titlelab="Global Population Structure, K=7",
  titlesize = 20)
```

```{r designate_cluster, message=F}
## extract one q file that reflects known species identity
global_k4_q <- slist[[41]]
global_k4_q$sampID <- row.names(global_k4_q)

## which cluster is carinulata?
gr_samples <- masterpopmap[grep( "CR", masterpopmap[,2]), ]
gr_q <- semi_join(global_k4_q, gr_samples)
max_cl_gr <- sapply(gr_q[,1:4], max)
elongata_cluster <- which(colnames(gr_q) == colnames(gr_q[which(max_cl_gr==1)]))
elongata_cluster

## which cluster is elongata?
ch_samples <- masterpopmap[grep( "CH", masterpopmap[,2]), ]
ch_q <- semi_join(global_k4_q, ch_samples)
max_cl_ch <- sapply(ch_q[,1:4], max)
carinulata_cluster <- which(colnames(ch_q) == colnames(ch_q[which(max_cl_ch==1)]))
carinulata_cluster

## which cluster is sublineata?
sub_samples <- masterpopmap[grep( "SUB_LAB", masterpopmap[,2]), ]
sub_q <- semi_join(global_k4_q, sub_samples)
max_cl_sub <- sapply(sub_q[,1:4], max)
sublineata_cluster <- which(colnames(sub_q) == colnames(sub_q[which(max_cl_sub==1)]))

## which cluster is carinata?
carina_samples <- masterpopmap[grep( "CARINA_LAB", masterpopmap[,2]),]
carina_q <- semi_join(global_k4_q, carina_samples)
max_cl_cara <- sapply(carina_q[,1:4], max)
carinata_cluster <- which(colnames(carina_q) == colnames(carina_q[which(max_cl_cara==1)]))

cls <- c(carinata_cluster, carinulata_cluster, sublineata_cluster, elongata_cluster)
spp <- c("carinata", "carinulata", "sublineata", "elongata")
clusters <- as.data.frame(cbind(spp = spp[order(cls)], cluster = seq(1,4)))


colnames(global_k4_q)[1:4] <- as.character(clusters$spp)
q_popmap <- merge(global_k4_q, popmap)
head(q_popmap)
```

```{r Global_K4_2save}
## slist[43]
## 1 = elongata
## 2 = carinata
## 3 = sub
## 4 = carinu
legend <- merge(legend_colors, clusters)
clustercols <- arrange(legend, cluster) %>%
  select(cbbPalette)
clustercols <- clustercols$cbbPalette

legendlab <- arrange(legend, cluster) %>%
  mutate(label = paste("D. ", spp))
legendlab <- legendlab$label[1:4]

plotQ(slist[41],
            exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/plots/",
        outputfilename =  "MainTextFigure_K4",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=clustercols,
  subsetgrp = pop_order,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  linesize=0.8,
  pointsize = 5,
  divtype = 1,
  grplabangle= 90,
  grplabpos = .9,
  grplabspacer = 0,
  grplabheight=35,
  grplabsize = 10,
  height = .05,
  width=100,
  showlegend = T, legendkeysize=30,legendtextsize=30, legendpos="left", legendlab=legendlab,
  # titlelab="Global Population Structure K=4",
  # titlesize = 20,
  basesize=30)
```

```{r plotmultiK}
# custom strip panel label showing k only
fn1 <- function(x) attr(x,"k")
spnames <- paste0("K=",sapply(slist,fn1))


# Combine K=2, K=3, K=4
k2k3k4 <- c(21,31,41)
plotQ(slist[k2k3k4],
            imgoutput="join",
            exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/plots/",
  outputfilename =  "Joined-K2-K3-K4",  
  exportplot=T,
  showyaxis=T,
  splab=spnames[k2k3k4],
  splabsize= 20,
  clustercol=clustercols,
  subsetgrp = pop_order,
  height = .1,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  grplabangle= 90,
  grplabpos = .8,
  grplabspacer = 0,
  grplabheight=70,
  width=100,
  showtitle=T,
  grplabsize = 8,
  # showlegend = T,
  # legendkeysize=20,legendtextsize=20, legendpos="right", legendlab=legendlab,
  titlelab="Global Population Structure across K",
  titlesize = 20)
```


```{r plotsubsets}
## Plot subsets of the whole set
subset1 <- sample_meta.dat %>% 
  arrange(Order) %>%
  select(popID) %>%
  slice(1:20)

subset2 <- sample_meta.dat %>% 
  arrange(Order) %>%
  select(popID) %>%
  slice(21:nrow(sample_meta.dat))

plotQ(slist[41],
            exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/plots/",
        outputfilename =  "subset1_global",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=clustercols,
  subsetgrp=as.character(subset1$popID),
  ordergrp = T,
  grplab = popID,
  linepos = .98,
  linesize=0.8,
  pointsize = 5,
  divtype = 1,
  grplabangle= 90,
  grplabpos = .9,
  grplabspacer = 0,
  grplabheight=55,
  grplabsize = 20,
  height = .05,
  width=100,
  showlegend = T, legendkeysize=30,legendtextsize=30, legendpos="left", legendlab=legendlab,
  basesize=30)

 
plotQ(slist[41],
            exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/structure_analysis/plots/",
        outputfilename =  "subset2_global",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=clustercols,
  subsetgrp=as.character(subset2$popID),
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  linesize=0.8,
  pointsize = 5,
  divtype = 1,
  grplabangle= 90,
  grplabpos = .9,
  grplabspacer = 0,
  grplabheight=45,
  grplabsize =  20,
  height = .05,
  width=100,
  basesize=30)

```


```{r}
## subset individuals for substructure of elongata 
elongata <- filter(global_k4_q, elongata > .1) %>%
  select('sampID')
nrow(elongata)
elongata_popmap <- semi_join(popmap, elongata)
write.table(elongata_popmap,
  file = "../info/elongata_popmap.tsv", 
            sep = "\t",
            quote=F, col.names = F, row.names = F)

## subset individuals for substructure of elongata 
carinulata <- filter(global_k4_q, carinulata > .1) %>%
  select('sampID')
nrow(carinulata)
carinulata_popmap <- semi_join(popmap, carinulata)
write.table(carinulata_popmap,
  file = "../info/carinulata_popmap.tsv", 
            sep = "\t",
            quote=F, col.names = F, row.names = F)

## Are population genetic statistics (like private alleles etc) to populations filtering?
others_popmap <- anti_join(popmap, carinulata)
write.table(others_popmap,
  file = "../info/others_popmap.tsv", 
            sep = "\t",
            quote=F, col.names = F, row.names = F)
```


## Which sites have individuals with admxiture? 
```{r designate_pops}
sample_size <- samplesize %>%
  select('popID', 'n')

num_admixed_inds <- q_popmap %>%
  filter_at(vars(elongata:sublineata),
            ~ . < .95) %>%
  group_by(popID) %>%
  count(popID)
colnames(num_admixed_inds)[2] <- 'num_admixed'

num_admixed_inds

merge(sample_size, num_admixed_inds) %>%
  mutate(perc_admixed = num_admixed/n) %>%
  arrange(desc(perc_admixed))

q_popmap %>%
  filter_at(vars(elongata:sublineata),
            ~ . < .95) %>%
  group_by(popID) %>%
  summarize_at(vars(carinata:sublineata), funs(mean, var))
```

## Which sites have multiple spp, even without admixture? 
```{r}
foo <- q_popmap %>%
  filter_all(any_vars(. > .95)) %>%
  group_by(popID) %>%
  summarize_at(vars(elongata:sublineata), funs(sum)) %>%
  merge(sample_size) 

max_ancestry <-  apply(X=foo[,2:5], MARGIN=1, FUN=max)
### if a cluster has less than foo$sample_size-.5, will return any pop with an ind less than 50% of majority ancestry 
mixed_pops <- foo[foo$n - .5 > max_ancestry,]
mixed_pops

perc_admixed <- merge(sample_size, num_admixed_inds) %>%
  mutate(perc_admixed = num_admixed/n) %>%
  arrange(desc(perc_admixed))
perc_admixed

mix_admix <- full_join(mixed_pops, perc_admixed[,c('popID', 'perc_admixed')]) %>%
  arrange(desc(perc_admixed)) 
mix_admix
```

There may be almost no population that is completely pure. How many 'pure' sites did we detect? 

```{r}
setdiff(popmap$popID, mix_admix$popID )
```


Among individuals, how many pure? How many hybrids, and of what type? 

```{r}
ci <- data.frame(attributes(slist[[41]])$ci)
colnames(ci) <- paste0(rep(colnames(global_k4_q)[1:4], each=2),
       "_", rep(c("L", "H"), times = 4))
ci <- cbind(ci, global_k4_q)
ci <- merge(ci, masterpopmap)
ci
write.csv(ci,"ancestry_confidenceintervals.csv",
          quote = FALSE)

library(tibble)  # for `rownames_to_column` and `column_to_rownames`
colnames(ci)

## How many tri-specific hybrids? 
ci %>%
  rownames_to_column('individual') %>%
  filter(carinata_L >.01 & elongata_L> .01 & sublineata_L > .01) %>%
  arrange(desc(elongata_L))

ci %>%
  rownames_to_column('individual') %>%
  filter(carinata >.01 & elongata > .01 & sublineata > .01) %>%
  arrange(desc(elongata))

## How many hybrids with carinulata?
ci %>%
  rownames_to_column('individual') %>%
  filter(carinulata_L >.01) %>%
  filter(elongata_L> .01 | 
           sublineata_L > .01 | carinata_L > .01) %>%
  arrange(desc(carinulata_L))

ci %>%
  rownames_to_column('individual') %>%
  filter(carinulata_H >.1) %>%
  filter(elongata_L> .01 | 
           sublineata_L > .01 | carinata_L > .01) %>%
  arrange(desc(carinulata_L))

## How many hybrids with sublineata?
ci %>%
  rownames_to_column('individual') %>%
  filter(sublineata_L >.01) %>%
  filter(carinulata_L > .01 | 
           elongata_L > .01 | carinata_L > .01) %>%
  arrange(desc(sublineata_L))


## How many carinata-sublineata hybrids? 
car_sub_hyb <- ci %>%
  rownames_to_column('individual') %>%
  filter(carinata_L >.01  & sublineata_L > .01 & elongata_L < .01)
car_sub_hyb

ci %>%
  rownames_to_column('individual') %>%
  filter(carinata_L >.01  & sublineata_L > .01 & elongata_L < .01) %>%
  count()

mean(car_sub_hyb$sublineata)

## How many carinata-elongata hybrids? 
car_elong_hyb <- ci %>%
  rownames_to_column('individual') %>%
  filter(carinata_L >.01 & elongata_L> .01 & sublineata_L < .01)

ci %>%
  rownames_to_column('individual') %>%
  filter(carinata_L >.01 & elongata_L> .01 & sublineata_L < .01) %>%
  count()

ci %>%
  rownames_to_column('individual') %>%
  filter(carinata_L >.01 & elongata_L> .01 & sublineata_L < .01) %>%
  summarize_at(vars(carinata_L, elongata_L), funs(mean))


## How many elong-sublineata hybrids? 
elong_sub_hyb <- ci %>%
  rownames_to_column('individual') %>%
  filter(carinata_L < .01  & sublineata_L > .01 & elongata_L > .01)
elong_sub_hyb
mean(elong_sub_hyb$sublineata)



```

```{r}
### this is probably better represented as a simplex (triangle)
pCS <- ggplot(car_sub_hyb) +
  geom_histogram(aes(carinata), 
                 alpha = .5, fill = legend[which(legend$spp=="carinata"), "cbbPalette"], stat = "bin", binwidth = .05) +
  geom_histogram(aes(sublineata), 
                 alpha = .5, fill = legend[which(legend$spp=="sublineata"), "cbbPalette"], stat = "bin", binwidth = .05) +
  ylim(c(0,8)) +
  scale_x_continuous(name= expression( ~italic("D. carinata")~ "X"  ~italic("D. sublineata")~ "q-values")) +
  theme_bw()

pCE <-ggplot(car_elong_hyb) +
  geom_histogram(aes(carinata), 
                 alpha = .5, fill = legend[which(legend$spp=="carinata"), "cbbPalette"], stat = "bin", binwidth = .05) +
  geom_histogram(aes(elongata), 
                 alpha = .5, fill = legend[which(legend$spp=="elongata"), "cbbPalette"], stat = "bin", binwidth = .05) +
  ylim(c(0,8)) +
    scale_x_continuous(name= expression( ~italic("D. elongata")~ "X"  ~italic(" D. carinata")~ "q-values")) +
  theme_bw()
pCE

pES <-ggplot(elong_sub_hyb) +
  geom_histogram(aes(elongata), 
                 alpha = .5,fill = legend[which(legend$spp=="elongata"), "cbbPalette"], stat = "bin", binwidth = .05) +
  geom_histogram(aes(sublineata), 
                 alpha = .5, fill = legend[which(legend$spp=="sublineata"), "cbbPalette"], stat = "bin", binwidth = .05) +
  ylim(c(0,8)) +
      scale_x_continuous(name= expression( ~italic("D. elongata")~ "X"  ~italic(" D. sublineata")~ "q-values")) +
  theme_bw()
  
cowplot::plot_grid(pCS, pCE, pES, nrow = 1, labels = c('A','B','C'))
ggsave("hybrid_distributions.jpg", dpi = 300, height = 4, width = 10, units = "in")

```

```{r}
## Pull pure parental species 
pure_parentals <- ci %>%
  filter(carinata_L > .995 | elongata_L > .995 | 
           sublineata_L > .995 | carinulata_L > .995) %>% 
  merge(popmap)

pure_parentals$assignment <- colnames(pure_parentals[,c('elongata', 'carinata', 'carinulata', 'sublineata')])[max.col(pure_parentals[,c('elongata', 'carinata', 'carinulata', 'sublineata')])]

table(pure_parentals$assignment)

pure_parentals_popmap <- pure_parentals[,c('sampID', 'assignment')]
write.table(pure_parentals_popmap, "../info/pure_parentals_popmap.tsv", col.names = F, quote = F, row.names = F, sep = "\t")
```



### Substructure

## Population structure within D. carinulata

```{r substructure_carinu}
sfiles <- list.files(path="~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/substructure/carinulata/structure_analysis/Results/", 
  pattern=glob2rx("outfile_*f"),full.names=TRUE)

carinu_slist <- readQ(sfiles,filetype="structure", 
               indlabfromfile = TRUE, 
               readci = TRUE)

sr1 <- summariseQ(tabulateQ(carinu_slist))
p <- evannoMethodStructure(data=sr1,exportplot=F,returnplot=T,returndata=F,basesize=12,linesize=0.7)
grid.arrange(p)

evannoMethodStructure(data=sr1,
                  exportplot=T,
                  exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/substructure/carinulata/structure_analysis/plots/",
                outputfilename = "carinulata_deltaK",
                returndata=F,
                height = 10, width = 9)

popmap <- left_join(data.frame(sampID = rownames(carinu_slist[[1]])), 
                    masterpopmap)

popID <- data.frame(popID = popmap$popID)
popID <- data.frame(lapply(popID, as.character), stringsAsFactors=FALSE)
sapply(popID, is.character)

pop_order <- sample_meta.dat[,c("popID", "Order")]
pop_order <- semi_join(pop_order, popID) %>%
  arrange(Order)
pop_order <- pop_order[,1]

legend_colors
carinu_legend <- legend_colors[c(4,6,7),]
carinu_cbbPalette <- carinu_legend[,'cbbPalette']

plotQ(alignK(carinu_slist[c(21:30)]),
            imgoutput="join",
                  exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/substructure/carinulata/structure_analysis/plots/",
  outputfilename =  "Joined-K2",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=carinu_cbbPalette,
  subsetgrp = pop_order,
  height = .1,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  grplabangle=45,
  grplabpos = .8,
  grplabspacer = 0,
  grplabheight=70,
  width=100,
  showtitle=T,
  grplabsize = 4,
  titlelab="D. carinulata Population Structure, K=2",
  titlesize = 20)


plotQ(alignK(carinu_slist[c(31:40)]),
            imgoutput="join",
                  exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/substructure/carinulata/structure_analysis/plots/",
  outputfilename =  "Joined-K3",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=carinu_cbbPalette,
  subsetgrp = pop_order,
  height = .1,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  grplabangle=45,
  grplabpos = .8,
  grplabspacer = 0,
  grplabheight=70,
  width=100,
  showtitle=T,
  grplabsize = 4,
  titlelab="D. carinulata Population Structure, K=3",
  titlesize = 20)

k4 <- plotQ(alignK(carinu_slist[c(41:50)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            subsetgrp = pop_order,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k4$plot[[1]])

k5 <- plotQ(alignK(carinu_slist[c(51:60)]),
            imgoutput="join",
            subsetgrp = pop_order,
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k5$plot[[1]])

k6 <- plotQ(alignK(carinu_slist[c(61:70)]),
            imgoutput="join",
            subsetgrp = pop_order,
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k6$plot[[1]])

k7 <- plotQ(alignK(carinu_slist[c(71:80)]),
            imgoutput="join",
            subsetgrp = pop_order,
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k7$plot[[1]])
```

```{r carinu_K3_2save}

plotQ(carinu_slist[31],
            exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/substructure/carinulata/structure_analysis/plots/",
        outputfilename =  "MainTextFigure_carinuK3",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=carinu_cbbPalette,
  subsetgrp = pop_order,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  linesize=0.8,
  pointsize = 5,
  divtype = 1,
  divsize = 1,
  grplabangle= 90,
  grplabpos = .9,
  grplabspacer = 0,
  grplabheight=40,
  grplabsize = 25,
  height = .05,
  width=100,
  showlegend = T, legendkeysize=30,legendtextsize=30, legendpos="left", legendlab= c("D. carinulata Fukang Ecotype", "D. carinulata Chilik Ecotype", "Other"),
  basesize=30)

```

## Population structure within D. elongata

```{r elongata_substructure}
sfiles <- list.files(path="~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/substructure/elongata/structure_analysis/Results/", 
  pattern="outfile_*",full.names=TRUE)

elong_slist <- readQ(sfiles,filetype="structure", 
               indlabfromfile = TRUE, 
               readci = TRUE)

popmap <- left_join(data.frame(sampID = rownames(elong_slist[[1]])), 
                    masterpopmap)

popID <- data.frame(popID = popmap$popID)
popID <- data.frame(lapply(popID, as.character), stringsAsFactors=FALSE)
sapply(popID, is.character)

pop_order <- sample_meta.dat[,c("popID", "Order")]
pop_order <- semi_join(pop_order, popID) %>%
  arrange(Order)
pop_order <- pop_order[,1]

legend_colors
elong_legend <- legend_colors[c(2,5,7),]
elong_cbbPalette <- elong_legend[,'cbbPalette']

```

```{r}
sr1 <- summariseQ(tabulateQ(elong_slist))

evannoMethodStructure(data=sr1,
                  exportplot=T,
                  exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/substructure/elongata/structure_analysis/plots/",
                outputfilename = "elongata_deltaK",
                returndata=F,
                height = 10, width = 9)

k2 <- plotQ(alignK(slist[c(21:30)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            subsetgrp = pop_order,
            ordergrp = T,
            grplab = popID,
            linepos = 1,
            grplabangle=-45,
            grplabpos = 0, 
            grplabheight=20)
grid.arrange(k2$plot[[1]])

k3 <- plotQ(alignK(slist[c(31:40)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k3$plot[[1]])

k4 <- plotQ(alignK(slist[c(41:50)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k4$plot[[1]])

k5 <- plotQ(alignK(slist[c(51:60)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k5$plot[[1]])

k6 <- plotQ(alignK(slist[c(61:70)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k6$plot[[1]])

k7 <- plotQ(alignK(slist[c(71:80)]),
            imgoutput="join",
            returnplot=T,exportplot=F,quiet=T,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k7$plot[[1]])
```

```{r elongata_substructure_K32Save}

plotQ(elong_slist[32],
            exportpath = "~/Code/diorhabda/StahlkeBitume_heirarchical_str/Git_Repo/substructure/elongata/structure_analysis/plots/",
        outputfilename =  "MainTextFigure_elongataK3",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=elong_cbbPalette,
  subsetgrp = pop_order,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  linesize=0.8,
  pointsize = 5,
  divtype = 1,
  divsize = 1,
  grplabangle= 90,
  grplabpos = .9,
  grplabspacer = 0,
  grplabheight=40,
  grplabsize = 25,
  height = .05,
  width=100,
  showlegend = T, legendkeysize=30,legendtextsize=30, legendpos="left", legendlab= c("D. elongata Mainland Ecotype", "D. elongata Crete Ecotype", "Other"),
  basesize=30)
```

