---
title: "structure_analysis"
author: "Amanda Stahlke"
date: "2/28/2020"
output: html_document
---

```{r setup, include=FALSE}
# devtools::install_github('royfrancis/pophelper')
library(pophelper)
library(gridExtra)
library(gtable)
library(label.switching)

library(dplyr)

getwd()

cbbPalette <- c("#56B4E9", "#E69F00", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#9370DB") 
human_colors <- c('blue', 'orange', 'green', 'yellow', 'dark blue', 'red', 'purple')
spp <- c('carinata', 'elongata', 'sublineata', 'carinulata', 'elongata_sub', 'carinulata_sub', 'other')

legend_colors <- as.data.frame(cbind(cbbPalette, human_colors, spp), stringsAsFactors = F)
legend_colors

masterpopmap <- read.delim("../info/masterpopmap.tsv",
                  sep = "\t",
                   col.names = c("sampID", "popID"),
                   stringsAsFactors = FALSE, header = FALSE)
table(masterpopmap$popID)

sample_meta.dat <- read.csv("../info/pop_metadata.csv", header = TRUE, stringsAsFactors = FALSE)
sample_meta.dat
```

# Structure Analysis

Examined "global" ancestry for all available samples. Old replicates and some suspected contamination/switching have been removed. The masterpopmap has ALL of these samples.

Admixture model, LD=0, allele freqs are uncorrelated, alpha is inferred for each popualtion.

Read in Results and set up popmap/poporder for visualization and downstream analysis. 
```{r global_structure_inputs}
sfiles <- list.files(path="Results/",
  pattern=glob2rx("outfile_*f"),full.names=TRUE)

slist <- readQ(sfiles,filetype="structure",
               indlabfromfile = TRUE,
               readci = TRUE)

popmap <- left_join(data.frame(sampID = rownames(slist[[1]])),
                    masterpopmap)

popID <- data.frame(popID = popmap$popID)
popID <- data.frame(lapply(popID, as.character), stringsAsFactors=FALSE)

pop_order <- sample_meta.dat[,c("popID", "Order")]
pop_order <- arrange(pop_order, Order)[,1]
```

Sample size for just Structure analysis

```{r}
samplesize <- left_join(popmap, sample_meta.dat) %>%
  count(popID) %>%
  merge(sample_meta.dat)
samplesize

write.table(samplesize, "pop_metadata_samplesize.tsv", sep = "\t", row.names = F, quote = F)

```

## What is the best K?

```{r which_K}
sr1 <- summariseQ(tabulateQ(slist))
p <- evannoMethodStructure(data=sr1,exportplot=F,returnplot=T,returndata=F,basesize=12,linesize=0.7)
grid.arrange(p)

evannoMethodStructure(data=sr1,
                  exportplot=T,
                  exportpath = "plots",
                outputfilename = "global_deltaK",
                returndata=F,
                height = 10, width = 9, 
                xaxisbreaks=c(1:nrow(sr1)))
```

K=2 and K=3 are super stable. D. carinata appears as a mix between sublineata and elongata. I am surprised to see, for the first time in these samples, indications of admixture between carinlata and sublineata along the Rio Grande.

As usual, K=4 reflects lab populations, elongata in the native range, and hints at carinulata substructure.

```{r alignK4-8}

k4 <- plotQ(alignK(slist[c(41:50)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            grplab = popID,
            ordergrp = T,
            grplabangle=-90)
grid.arrange(k4$plot[[1]])

```

Code to save a nice version of all reps of K=4 for the manuscript. 

```{r plots2save_repsK}

fn1 <- function(x) attr(x,"k")
spnames <- rep("", 10)

plotQ(alignK(slist[c(41:50)]),
            imgoutput="join",
            exportpath = "plots",
  outputfilename =  "Joined-K4",  
  exportplot=T,
  showyaxis=F,
  showsp = T,
  splab=rep("", 10),
  clustercol=cbbPalette,
  subsetgrp = pop_order,
  height = .005,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  grplabangle=90,
  grplabpos = .8,
  grplabspacer = 0,
  basesize=14,
  grplabsize = 15,
  width = 150,
  grplabheight=150)
```

## Assign species identity to each cluster
Extract one q file that reflects known species identity

```{r designate_cluster, message=F}
global_k4_q <- slist[[41]]
global_k4_q$sampID <- row.names(global_k4_q)

## which cluster is carinulata?
gr_samples <- masterpopmap[grep( "CR", masterpopmap[,2]), ]
gr_q <- semi_join(global_k4_q, gr_samples)
max_cl_gr <- sapply(gr_q[,1:4], max)
elongata_cluster <- which(colnames(gr_q) == colnames(gr_q[which(max_cl_gr==1)]))
elongata_cluster

## which cluster is elongata?
ch_samples <- masterpopmap[grep( "CH", masterpopmap[,2]), ]
ch_q <- semi_join(global_k4_q, ch_samples)
max_cl_ch <- sapply(ch_q[,1:4], max)
carinulata_cluster <- which(colnames(ch_q) == colnames(ch_q[which(max_cl_ch==1)]))
carinulata_cluster

## which cluster is sublineata?
sub_samples <- masterpopmap[grep( "SUB_LAB", masterpopmap[,2]), ]
sub_q <- semi_join(global_k4_q, sub_samples)
max_cl_sub <- sapply(sub_q[,1:4], max)
sublineata_cluster <- which(colnames(sub_q) == colnames(sub_q[which(max_cl_sub==1)]))

## which cluster is carinata?
carina_samples <- masterpopmap[grep( "CARINA_LAB", masterpopmap[,2]),]
carina_q <- semi_join(global_k4_q, carina_samples)
max_cl_cara <- sapply(carina_q[,1:4], max)
carinata_cluster <- which(colnames(carina_q) == colnames(carina_q[which(max_cl_cara==1)]))

cls <- c(carinata_cluster, carinulata_cluster, sublineata_cluster, elongata_cluster)
spp <- c("carinata", "carinulata", "sublineata", "elongata")
clusters <- as.data.frame(cbind(spp = spp[order(cls)], cluster = seq(1,4)))

colnames(global_k4_q)[1:4] <- as.character(clusters$spp)
q_popmap <- merge(global_k4_q, popmap)
head(q_popmap)
```

```{r}
legend <- merge(legend_colors, clusters)
clustercols <- arrange(legend, cluster) %>%
  select(cbbPalette)
clustercols <- clustercols$cbbPalette

legendlab <- arrange(legend, cluster) %>%
  mutate(label = paste("D. ", spp))
legendlab <- legendlab$label[1:4]
```

Code to save a nice version of one rep of K=4 for the manuscript. 

```{r Global_K4_2save, message=F}
plotQ(slist[41],
            exportpath = "plots",
        outputfilename =  "MainTextFigure_K4",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=clustercols,
  subsetgrp = pop_order,
  sortind="all",
  grplab = popID,
  linepos = 1,
  linesize=0.8,
  pointsize = 5,
  divtype = 1,
  divsize=1.1,
  grplabangle= 90,
  grplabpos = .9,
  grplabspacer = 0,
  grplabheight=35,
  grplabsize = 10,
  height = .05,
  width=100,
  showlegend = T, legendkeysize=30,legendtextsize=30, legendpos="left", legendlab=legendlab,
  # titlelab="Global Population Structure K=4",
  # titlesize = 20,
  basesize=30)
```

Code to save a nice version of K=2-4 for the manuscript. 

```{r k2_k3_k4_2save, message=F}
# custom strip panel label showing k only
fn1 <- function(x) attr(x,"k")
spnames <- paste0("K=",sapply(slist,fn1))

# Combine K=2, K=3, K=4
k2k3k4 <- c(21,31,41)
plotQ(slist[k2k3k4],
            imgoutput="join",
            exportpath = "plots",
  outputfilename =  "Joined-K2-K3-K4",  
  exportplot=T,
  showyaxis=F,
  clustercol=clustercols,
  subsetgrp = pop_order,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  # divsize = 1.2,
  grplabangle= 90,
  grplabpos = .8,
  grplabspacer = 0,
  basesize=14,
  showsp = F,
  grplabsize = 6,
  width = 30,
  grplabheight=40)
```

## Exploratory visualizations that are not presented in the manuscript

Multi-line K=4 to visulize individuals
```{r multiline_inds_K4}
plotQMultiline(slist[41],
      showticks=T,
      clustercol=cbbPalette,
      grplab = popID,
      ordergrp=TRUE,
      spl = 25,
      showindlab = T,
      showyaxis = T,
      barsize = .9,
      subsetgrp = pop_order,
      basesize=11,
      exportpath = "plots",
      outputfilename = "multilineK4",
      useindlab = T)
```

All reps of each run to visualize results across K values
```{r Ks2-8}
k2 <- plotQ(alignK(slist[c(21:30)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            subsetgrp = pop_order,
            ordergrp = T,
            grplab = popID,
            linepos = 1,
            grplabangle=-45,
            grplabpos = 0,
            grplabheight=20)
grid.arrange(k2$plot[[1]])

k3 <- plotQ(alignK(slist[c(31:40)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k3$plot[[1]])


k5 <- plotQ(alignK(slist[c(51:60)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            grplab = popID,
            ordergrp = T,
            grplabangle=-90)
grid.arrange(k5$plot[[1]])

k6 <- plotQ(alignK(slist[c(61:70)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            grplab = popID,
            ordergrp = T,
            grplabangle=-90)
grid.arrange(k6$plot[[1]])


k7 <- plotQ(alignK(slist[c(71:80)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k7$plot[[1]])

k8 <- plotQ(alignK(slist[c(81:90)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            # clustercol=cbbPalette,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k8$plot[[1]])

```

## Among individuals, how many pure? How many hybrids, and of what type?

Extract confidence intervals. 

```{r}
ci <- data.frame(attributes(slist[[41]])$ci)
colnames(ci) <- paste0(rep(colnames(global_k4_q)[1:4], each=2),
       "_", rep(c("L", "H"), times = 4))
ci <- cbind(ci, global_k4_q)
ci <- merge(ci, masterpopmap)
head(ci)

write.csv(ci,"ancestry_confidenceintervals.csv",
          quote = FALSE)

```

Use "diagnostic" samples to define threshold. 
```{r}
diag_sublineata <- c("SUB_LAB", "47TX", "49TX", "50TX", "51TX", "52TX")
diagnostic <- data.frame(popID = diag_sublineata, spp=rep("sublineata", length(diag_sublineata)))
diag_elongata <- c("43GR", "44GR", "37CR", "38CR", "39CR", "41CR")
diagnostic <- rbind(diagnostic, 
      data.frame(popID = diag_elongata, spp=rep("elongata", length(diag_elongata))))

ci %>%
  semi_join(diagnostic) %>%
  arrange(desc(carinulata_L))

find_threshold <- ci %>%
  semi_join(diagnostic) %>%
  arrange(desc(carinulata_L))

t <- find_threshold[1,"carinulata"]
t

t_low <- find_threshold[1,"carinulata_L"]
t_low
```
### Each hybrid type

How many hybrids with carinulata?
```{r carinulata}
carinu_hyb <- ci %>%
  dplyr::filter(carinulata_L > t_low ) %>%
  dplyr::filter(elongata_L > t_low |
           sublineata_L > t_low | 
             carinata_L > t_low) %>%
  arrange(desc(carinulata_L))

carinu_hyb
```


How many tri-specific hybrids?
```{r tri-specific}
ci %>%
  dplyr::filter(carinata_L > t_low & elongata_L > t_low & sublineata_L > t_low) 

## If we count less-conservatively:
ci %>%
  dplyr::filter(carinata_L >.01 & elongata_L> .01 & sublineata_L > .01) %>%
  arrange(desc(elongata_L))

```


How many hybrids with sublineata?
```{r sublineata}
ci %>%
  dplyr::filter(sublineata_L > t_low) %>%
  dplyr::filter(carinulata_L > t_low |
           elongata_L > t_low | carinata_L > t_low) %>%
  arrange(desc(sublineata_L)) 

ci %>%
  dplyr::filter(sublineata_L > t) %>%
  dplyr::filter(carinulata_L > t_low |
           elongata_L > t_low | carinata_L > t_low) %>%
  arrange(desc(sublineata_L)) %>%
  count()
```

D. elongata hybrids

```{r elongata}
ci %>%
  dplyr::filter(elongata_L > t_low) %>%
  dplyr::filter(carinulata_L > t_low |
           sublineata_L > t_low | carinata_L > t_low) %>%
  arrange(desc(elongata_L))

ci %>%
  dplyr::filter(elongata_L > t_low) %>%
  dplyr::filter(carinulata_L > t_low |
           sublineata_L > t_low | carinata_L > t_low) %>%
  arrange(desc(elongata_L)) %>%
  count()
```

### Specific bi-parental types
How many carinata-sublineata hybrids?
How many carinata-elongata hybrids?
How many elong-sublineata hybrids?

```{r  bi-parental}
car_sub_hyb <- ci %>%
  dplyr::filter(carinata_L >t_low  & sublineata_L > t_low & elongata_L < t_low)
car_sub_hyb

car_sub_hyb %>%
  summarise(mean(sublineata), nrow(.))

car_elong_hyb <- ci %>%
  dplyr::filter(carinata_L > t_low & elongata_L> t_low & sublineata_L < t_low) %>%
  arrange(desc(elongata_L))
car_elong_hyb
car_elong_hyb %>%
  summarise(mean(carinata), nrow(.))


elong_sub_hyb <- ci %>%
  dplyr::filter(carinata_L < t_low  & sublineata_L > t_low & elongata_L > t_low)
elong_sub_hyb

elong_sub_hyb %>%
  summarise(mean(sublineata), nrow(.))
```

## Vizualize bi-parental crosses!
This could be better represented as a simplex (triangle)?

```{r viz_bi-parental}
pCS <- ggplot(car_sub_hyb) +
  geom_histogram(aes(carinata),
                 alpha = .5, fill = legend[which(legend$spp=="carinata"), "cbbPalette"], stat = "bin", binwidth = .05) +
  geom_histogram(aes(sublineata),
                 alpha = .5, fill = legend[which(legend$spp=="sublineata"), "cbbPalette"], stat = "bin", binwidth = .05) +
  ylim(c(0,8)) +
  scale_x_continuous(name= expression( ~italic("D. carinata")~ "X"  ~italic("D. sublineata")~ "q-values")) +
  theme_bw()
pCS

pCE <-ggplot(car_elong_hyb) +
  geom_histogram(aes(carinata),
                 alpha = .5, fill = legend[which(legend$spp=="carinata"), "cbbPalette"], stat = "bin", binwidth = .05) +
  geom_histogram(aes(elongata),
                 alpha = .5, fill = legend[which(legend$spp=="elongata"), "cbbPalette"], stat = "bin", binwidth = .05) +
  ylim(c(0,8)) +
    scale_x_continuous(name= expression( ~italic("D. elongata")~ "X"  ~italic(" D. carinata")~ "q-values")) +
  theme_bw()
pCE

pES <-ggplot(elong_sub_hyb) +
  geom_histogram(aes(elongata),
                 alpha = .5,fill = legend[which(legend$spp=="elongata"), "cbbPalette"], stat = "bin", binwidth = .05) +
  geom_histogram(aes(sublineata),
                 alpha = .5, fill = legend[which(legend$spp=="sublineata"), "cbbPalette"], stat = "bin", binwidth = .05) +
  ylim(c(0,8)) +
      scale_x_continuous(name= expression( ~italic("D. elongata")~ "X"  ~italic(" D. sublineata")~ "q-values")) +
  theme_bw()

cowplot::plot_grid(pCS, pCE, pES, nrow = 1, labels = c('A','B','C'))
ggsave("hybrid_distributions.jpg", dpi = 300, height = 4, width = 10, units = "in")

```


## Which sites have individuals with admxiture?
```{r designate_pops}
sample_size <- samplesize %>%
  select('popID', 'n')

hybrids <- rbind(elong_sub_hyb, car_elong_hyb, carinu_hyb, car_sub_hyb)
nrow(hybrids)

no_hybrids <- anti_join(ci, hybrids)
pure_carinu <- dplyr::filter(no_hybrids, carinulata > .9)  
 nrow(pure_carinu)
pure_sub <- dplyr::filter(no_hybrids, sublineata > .9)  
   nrow(pure_sub)
pure_carinata <- dplyr::filter(no_hybrids, carinata > .9 )  
   nrow(pure_carinata)
pure_elongata <- dplyr::filter(no_hybrids, elongata > .9 )  
   nrow(pure_elongata)   

pure_parentals <- rbind(pure_carinu, pure_sub, pure_carinata, pure_elongata)

what_else <- anti_join( ci, pure_parentals) %>%
  anti_join(., hybrids)
  
 num_admixed_inds <- hybrids %>%
  group_by(popID) %>%
  count(popID)
colnames(num_admixed_inds)[2] <- 'num_admixed'

num_admixed_inds

merge(sample_size, num_admixed_inds) %>%
  mutate(perc_admixed = num_admixed/n) %>%
  arrange(desc(perc_admixed))

hybrids %>%
  filter_at(vars(elongata:carinata),
            ~ . < .95) %>%
  group_by(popID) %>%
  summarize_at(vars(carinata:sublineata), funs(mean, var))
```

## Which sites have multiple spp, even without admixture?
```{r}
foo <- q_popmap %>%
  filter_all(any_vars(. > .95)) %>%
  group_by(popID) %>%
  summarize_at(vars(elongata:carinata), funs(sum)) %>%
  merge(sample_size)

max_ancestry <-  apply(X=foo[,2:5], MARGIN=1, FUN=max)
### if a cluster has less than foo$sample_size-.5, will return any pop with an ind less than 50% of majority ancestry
mixed_pops <- foo[foo$n - .5 > max_ancestry,]
mixed_pops

perc_admixed <- merge(sample_size, num_admixed_inds) %>%
  mutate(perc_admixed = num_admixed/n) %>%
  arrange(desc(perc_admixed))
perc_admixed

mix_admix <- full_join(mixed_pops, perc_admixed[,c('popID', 'perc_admixed')]) %>%
  arrange(desc(perc_admixed))
mix_admix
```

How many 'pure' sites did we detect?

```{r}
setdiff(popmap$popID, mix_admix$popID )
length(setdiff(popmap$popID, mix_admix$popID ))
```

Pull pure parental species for reference down the road. 

```{r}
pure_parentals <- ci %>%
  dplyr::filter(carinata_L > 1 - t_low | elongata_L > 1 - t_low |
           sublineata_L > 1 - t_low | carinulata_L > 1 - t_low) %>%
  merge(popmap)

pure_parentals$assignment <- colnames(pure_parentals[,c('elongata', 'carinata', 'carinulata', 'sublineata')])[max.col(pure_parentals[,c('elongata', 'carinata', 'carinulata', 'sublineata')])]

table(pure_parentals$assignment)

pure_parentals_popmap <- pure_parentals[,c('sampID', 'assignment')]
write.table(pure_parentals_popmap, "../info/pure_parentals_popmap.tsv", col.names = F, quote = F, row.names = F, sep = "\t")
```

# Substructure
Subset individuals for substructure of elongata, carinulata, and everything but carinulata ("others").

```{r}
elongata <- dplyr::filter(ci, elongata_L > t_low) %>%
  select('sampID', 'popID')
nrow(elongata)

write.table(elongata,
  file = "../info/elongata_popmap.tsv",
            sep = "\t",
            quote=F, col.names = F, row.names = F)

carinulata <- dplyr::filter(ci, carinulata_L > t_low) %>%
  select('sampID', 'popID')
nrow(carinulata)

write.table(carinulata,
  file = "../info/carinulata_popmap.tsv",
            sep = "\t",
            quote=F, col.names = F, row.names = F)

others_popmap <- anti_join(popmap, carinulata)
write.table(others_popmap,
  file = "../info/others_popmap.tsv",
            sep = "\t",
            quote=F, col.names = F, row.names = F)
```

## Population structure within D. carinulata

```{r substructure_carinu_results}
sfiles <- list.files(path="../substructure/carinulata/structure_analysis/Results/",
  pattern=glob2rx("outfile_*f"),full.names=TRUE)

carinu_slist <- readQ(sfiles,filetype="structure",
               indlabfromfile = TRUE,
               readci = TRUE)

popmap <- left_join(data.frame(sampID = rownames(carinu_slist[[1]])),
                    masterpopmap)

popID <- data.frame(popID = popmap$popID)
popID <- data.frame(lapply(popID, as.character), stringsAsFactors=FALSE)
sapply(popID, is.character)

pop_order <- sample_meta.dat[,c("popID", "Order")]
pop_order <- semi_join(pop_order, popID) %>%
  arrange(Order)
pop_order <- pop_order[,1]

legend_colors
carinu_legend <- legend_colors[c(4,6,7),]
carinu_cbbPalette <- carinu_legend[,'cbbPalette']
carinu_cbbPalette
```


```{r substructure_carinu}
sr1 <- summariseQ(tabulateQ(carinu_slist))
p <- evannoMethodStructure(data=sr1,exportplot=F,returnplot=T,returndata=F,basesize=12,linesize=0.7)
grid.arrange(p)

evannoMethodStructure(data=sr1,
                  exportplot=T,
                  exportpath = "../substructure/carinulata/structure_analysis/plots/",
                outputfilename = "carinulata_deltaK",
                returndata=F,
                height = 10, width = 9, 
                xaxisbreaks=c(1:nrow(sr1)))


plotQ(alignK(carinu_slist[c(21:30)]),
            imgoutput="join",
                  exportpath = "../substructure/carinulata/structure_analysis/plots/",
  outputfilename =  "Joined-K2",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=carinu_cbbPalette,
  subsetgrp = pop_order,
  height = .1,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  grplabangle=45,
  grplabpos = .8,
  grplabspacer = 0,
  grplabheight=70,
  width=100,
  showtitle=T,
  grplabsize = 4,
  titlelab="D. carinulata Population Structure, K=2",
  titlesize = 20)


plotQ(alignK(carinu_slist[c(31:40)]),
            imgoutput="join",
                  exportpath = "../substructure/carinulata/structure_analysis/plots/",
  outputfilename =  "Joined-K3",  
  exportplot=T,
  showyaxis=T,
  showsp = F,
  clustercol=carinu_cbbPalette,
  subsetgrp = pop_order,
  height = .1,
  ordergrp = T,
  grplab = popID,
  linepos = 1,
  divtype = 1,
  grplabangle=45,
  grplabpos = .8,
  grplabspacer = 0,
  grplabheight=70,
  width=100,
  showtitle=T,
  grplabsize = 4,
  titlelab="D. carinulata Population Structure, K=3",
  titlesize = 20)

k4 <- plotQ(alignK(carinu_slist[c(41:50)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            subsetgrp = pop_order,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k4$plot[[1]])

k5 <- plotQ(alignK(carinu_slist[c(51:60)]),
            imgoutput="join",
            subsetgrp = pop_order,
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k5$plot[[1]])

k6 <- plotQ(alignK(carinu_slist[c(61:70)]),
            imgoutput="join",
            subsetgrp = pop_order,
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k6$plot[[1]])

k7 <- plotQ(alignK(carinu_slist[c(71:80)]),
            imgoutput="join",
            subsetgrp = pop_order,
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k7$plot[[1]])
```

```{r carinu_K3_2save}
plotQ(carinu_slist[32],
  exportpath = "../substructure/carinulata/structure_analysis/plots/",
  outputfilename =  "MainTextFigure_carinuK3",  
  exportplot=T,
  showyaxis=T,
  clustercol=carinu_cbbPalette,
  subsetgrp = pop_order,
  sortind="all",
  grplab = popID,
  linepos = 1,
  linesize=0.8,
  pointsize = 5,
  divtype = 1,
  divsize = 1.2,
  grplabangle= 90,
  grplabpos = .9,
  grplabspacer = 0,
  grplabheight=40,
  grplabsize = 25,
  height = .05,
  width=100,
  showlegend = T,
  showsp = F, 
  legendkeysize=40,legendtextsize=40, legendpos="left", legendlab= c("D. carinulata Fukang Ecotype", "D. carinulata Chilik Ecotype", "Other"),
  basesize=40)

```

## Population structure within D. elongata

```{r elongata_substructure}
sfiles <- list.files(path="../substructure/elongata/structure_analysis/Results/",
  pattern=glob2rx("outfile_*f"),full.names=TRUE)

elong_slist <- readQ(sfiles,filetype="structure",
               indlabfromfile = TRUE,
               readci = TRUE)

popmap <- left_join(data.frame(sampID = rownames(elong_slist[[1]])),
                    masterpopmap)

popID <- data.frame(popID = popmap$popID)
popID <- data.frame(lapply(popID, as.character), stringsAsFactors=FALSE)
sapply(popID, is.character)

pop_order <- sample_meta.dat[,c("popID", "Order")]
pop_order <- semi_join(pop_order, popID) %>%
  arrange(Order)
pop_order <- pop_order[,1]

legend_colors
elong_legend <- legend_colors[c(2,5,7),]
elong_cbbPalette <- elong_legend[,'cbbPalette']
elong_cbbPalette
```

```{r}
sr1 <- summariseQ(tabulateQ(elong_slist))

evannoMethodStructure(data=sr1,
                  exportplot=T,
                  exportpath = "../substructure/elongata/structure_analysis/plots/",
                outputfilename = "elongata_deltaK",
                returndata=F,
                height = 10, width = 9,
                xaxisbreaks=c(1:nrow(sr1)))

k2 <- plotQ(alignK(elong_slist[c(21:30)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            subsetgrp = pop_order,
            ordergrp = T,
            grplab = popID,
            linepos = 1,
            grplabangle=-45,
            grplabpos = 0,
            grplabheight=20)
grid.arrange(k2$plot[[1]])

k3 <- plotQ(alignK(elong_slist[c(31:40)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k3$plot[[1]])

k4 <- plotQ(alignK(elong_slist[c(41:50)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k4$plot[[1]])

k5 <- plotQ(alignK(elong_slist[c(51:60)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k5$plot[[1]])

k6 <- plotQ(alignK(elong_slist[c(61:70)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            clustercol=cbbPalette,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k6$plot[[1]])

k7 <- plotQ(alignK(elong_slist[c(71:80)]),
            imgoutput="join",
            returnplot=T,exportplot=F,basesize=11,
            ordergrp = T,
            grplab = popID,
            grplabangle=-90)
grid.arrange(k7$plot[[1]])
```

```{r elongata_substructure_K32Save}

plotQ(elong_slist[37],
            exportpath = "../substructure/elongata/structure_analysis/plots/",
        outputfilename =  "MainTextFigure_elongataK3",  
  exportplot=T,
  showyaxis=F,
  clustercol=elong_cbbPalette,
  subsetgrp = pop_order,
  sortind="all",
  grplab = popID,
  linepos = 1,
  linesize=0.8,
  pointsize = 5,
  divtype = 1,
  divsize = 1.2,
  grplabangle= 90,
  grplabpos = .9,
  grplabspacer = 0,
  grplabheight=40,
  grplabsize = 25,
  height = .05,
  width=100,
  showlegend = F,
  showsp = F,
  basesize=40)
```

# Is there a latitudinal gradient found for ecotypes or ancestry?

```{r}
library(ggtext) 

global_q.dat <- slist[[41]]
global_q.dat$sampID <- rownames(global_q.dat)
global_q.dat.popmap.meta <- merge(global_q.dat,
                    masterpopmap) %>%
  merge(sample_meta.dat)   %>%
 dplyr::filter(Grp == "potential_hybrid_zone")

sub_plot <- ggplot(global_q.dat.popmap.meta, aes(x = Latitude, y = Cluster2)) +
  geom_smooth(method = "lm", color = "red") +
  geom_jitter(aes(color = popID, shape = cat), alpha = .8, size = 4) +
  ylim(c(0,1)) +
  ylab("Proportion of *D. sublineata* ancestry") +
  guides(col = guide_legend(ncol = 2)) +
  theme_bw(base_size = 12)+
  theme(axis.title.y = element_markdown())
sub_plot
ggsave("sublineata_ancestry_latitude.jpg")

fit_carina <- lm(global_q.dat.popmap.meta$Cluster4 ~ global_q.dat.popmap.meta$Latitude)
summary(fit_carina)

fit_sub <- lm(global_q.dat.popmap.meta$Cluster2 ~ global_q.dat.popmap.meta$Latitude)
summary(fit_sub)

plot(fit_sub, which=1, col=c("blue"))
plot(fit_sub, which=2, col=c("blue"))

global_q.dat.popmap.meta$predicted <- predict(fit_sub)   # Save the predicted values
global_q.dat.popmap.meta$residuals <- residuals(fit_sub) # Save the residual values

ggplot(global_q.dat.popmap.meta, aes(x = Latitude, y = Cluster2)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +     # regression line  
  geom_segment(aes(xend = Latitude, yend = predicted), alpha = .2) +      # draw line from point to line
  geom_point(aes(color = abs(residuals), size = abs(residuals))) +  # size of the points
  scale_color_continuous(low = "green", high = "red") +             # colour of the points mapped to residual size - green smaller, red larger
  guides(color = FALSE, size = FALSE) +                             # Size legend removed
  geom_point(aes(y = predicted), shape = 1) +
  ylab("Proportion of D. sublineata ancestry") +
  theme_bw()
ggsave("sublineata_ancestry_latitude_fit.jpg")

```


```{r plot_q_v_latitude_carinu}
carinu_q.dat <- carinu_slist[[32]]
carinu_q.dat$sampID <- rownames(carinu_q.dat)
carinu_q.popmap.meta <- merge(carinu_q.dat,
                    masterpopmap) %>%
  merge(sample_meta.dat) #%>%
  # dplyr::filter(Cluster3 < .5)  # %>%
  # dplyr::filter( popID != "46CH"  & popID != "1WY"& popID != "18TX" &  popID != "28NM")


carinu_plot <- ggplot(carinu_q.popmap.meta) +
  geom_smooth(aes(Latitude, Cluster2), method = "lm", color = "red") +
  geom_jitter(aes(x = Latitude, y = Cluster2, color = popID, shape = cat), alpha = .8, size = 4) +
  ylab("Proportion of Chilik ancestry") +
  ylim(c(0,1)) +
  guides(col = guide_legend(ncol = 2)) +
  theme_bw(base_size = 12) 
carinu_plot
ggsave("fukang_ancestry_latitude.jpg")

fit_carinu <- lm(carinu_q.popmap.meta$Cluster2 ~ carinu_q.popmap.meta$Latitude)
summary(fit_carinu)
plot(fit_carinu, which=1, col=c("blue"))
plot(fit_carinu, which=2, col=c("blue"))

carinu_q.popmap.meta$predicted <- predict(fit_carinu)   # Save the predicted values
carinu_q.popmap.meta$residuals <- residuals(fit_carinu) # Save the residual values

ggplot(carinu_q.popmap.meta, aes(x = Latitude, y = Cluster2)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +     # regression line  
  geom_segment(aes(xend = Latitude, yend = predicted), alpha = .2) +      # draw line from point to line
  geom_point(aes(color = abs(residuals), size = abs(residuals))) +  # size of the points
  scale_color_continuous(low = "green", high = "red") +
  ylab("Proportion of Fukang ancestry") + # colour of the points mapped to residual size - green smaller, red larger
  guides(color = FALSE, size = FALSE) +      
    ylab("Proportion of Chilik ancestry") +# Size legend removed
  geom_point(aes(y = predicted), shape = 1) +
  theme_bw()
ggsave("fukang_ancestry_latitude_fit.jpg")
```

```{r}
library(cowplot)
plots <- align_plots( sub_plot, carinu_plot, align = 'v', axis = 'l')
plot_grid(plots[[1]],
          plots[[2]],
          ncol = 1, 
          labels = c('A', 'B'))
ggsave("sublineata_carinulata_ancestry_latitude.png", dpi = 300, height = 8, width = 8, units = "in")
```

